LHC abort gap monitor electronics upgrade

### 2024-05-06 paper.tex
Update on Overleaf.

\section{Introduction}
During the flat-top phase, the LHC ring stores up to \SI{350}{MJ} of beam energy,
spread across 2808 bunches with an average beam transverse size under
\SI{1}{mm}. This high energy density can melt metallic surfaces and quench
superconducting magnets, posing a risk to the machine~\cite{LHC_report}. 
To safely extract the beam from the vacuum compartments, a series of kicker
magnets belonging to the beam dump system is used. These magnets take \SI{3}{\micro\second} to
ramp up their magnetic field~\cite{beam_dump_system}. This defines the size of
an abort gap (AG) in the bunch train. Due to system imperfections, a residual amount of particles populate this gap.  The typical AG population levels are two to three orders of magnitude lower in comparison to a single LHC bunch.
When the beam dump is initiated, the particles in AG do not receive full
extraction force. That causes deflection towards the vacuum chamber. When a certain energy accumulates in the AG, magnets can quench. Thus,
a~continuous monitoring of the AG population is necessary, which is the purpose
of the AGM \cite{particles_in_ag}.\\
This paper reviews the upgrade of the current AGM system in the LHC since 2010,
focusing on system architecture, analog electronics chain, temperature
dependence, timing instabilities, and a comparison with the upgraded system.

\section{System architecture}
The AGM system comprises of three main parts. The first part is the optical front-end
(OF), which includes~a pre-amplifier positioned beneath the vacuum chamber. The
second part is a data acquisition system (DAQ) deployed on an FPGA platform
located outside the radiation area. The third part involves a~control software (SW) that
retrieves, interprets, and publishes the measured data, facilitating interaction
with the instrument. The system architecture is illustrated in
Fig.~\ref{fig:bsra_chain}, with each part labeled accordingly.
\begin{figure}[!tbh]
    \centering
    \includegraphics[width=\columnwidth]{bsra_chain.pdf}
    \caption{BSRA Chain.}
    \label{fig:bsra_chain}
\end{figure} 
The diagram on the left depicts the layout of the LHC, with beam 1 represented
in red and beam 2 in blue. It includes all four experiments (Alice, Atlas,
LHCb, and CMS). The AGM is located at Point 4 (indicated by the yellow
square), next to the CMS experiment. The Timing, Trigger, and Control system
(TTC) is shown as a purple rectangle in the center of the LHC ring. This system
receives RF clocking and broadcasts it into the network (TTC
distribution)~\cite{ttc_distribution}. The AGM extracts bunch clock information
(corresponding to the position of bunches) and turn clock information (marking
the start of a~turn) from the LHC timing block messages.\\
Beam intensity is measured using the synchrotron light (SL) radiation, as indicated by the red line directed towards the Multi-channel photomultiplier
(MCP-PMT) in the optical front-end (OF). The beam traverses first Neutral Density optical filters, which increases the system dynamic range. Subsequently, the signal is
transmitted over a \SI{100}{m}-long cable. To compensate for losses in the cable,
the signal is amplified beforehand. Then it proceeds to the Data Acquisition
(DAQ) system, which measures \SI{6.4}{\micro\second} window and integrates the
intervals of \SI{100}{\nano\second}. This creates set of 64 values published to
the SW every \SI{10}{\micro\second}.

>>>>

\section{Introduction}
During the flat-top phase, the LHC ring stores up to \SI{350}{MJ} of beam energy,
spread across 2808 bunches with an average beam transverse size under
\SI{1}{mm}. This high energy density can melt metallic surfaces and quench
superconducting magnets, posing a risk to the machine~\cite{LHC_report}. 
To safely extract the beam from the vacuum compartments, a series of kicker
magnets belonging to the beam dump system is used. These magnets take \SI{3}{\micro\second} to
ramp up their magnetic field~\cite{beam_dump_system}. This defines the size of
an abort gap (AG) in the bunch train. Due to system imperfections, a residual amount of particles populate this gap.  The typical AG population levels are two to three orders of magnitude lower in comparison to a single LHC bunch.
When the beam dump is initiated, the particles in AG do not receive full
extraction force. That causes the particles in the AG to deflect towards the vacuum chamber. When a certain energy accumulates in the AG, magnets can quench. Thus,
a~continuous monitoring of the AG population is necessary, which is the purpose
of the AGM \cite{particles_in_ag}.\\
This paper reviews the upgrade of the current AGM system in the LHC since 2010,
focusing on system architecture, analog electronics chain, temperature
dependence, timing instabilities, and a comparison with the upgraded system.

\section{System architecture}
The AGM system consists of three main parts:

\begin{itemize}
    \item \textbf{optical front-end (OF)}, which includes~a pre-amplifier located beneath the vacuum chamber,
    \item \textbf{data acquisition system (DAQ)} deployed on an FPGA platform located outside the radiation area
    \item \textbf{a~control software (SW)} that retrieves, interprets, and publishes the measured data, facilitating interaction with the instrument
\end{itemize}

The system architecture is illustrated in
Fig.~\ref{fig:bsra_chain}, with each part labelled accordingly.
\begin{figure}[!tbh]
    \centering
    \includegraphics[width=\columnwidth]{bsra_chain.pdf}
    \caption{BSRA Chain.}
    \label{fig:bsra_chain}
\end{figure} 
The drawing depicts the layout of the LHC, with beam 1 represented
in red and beam 2 in blue. It includes all four experiments (Alice, Atlas,
LHCb, and CMS). The AGM is located at Point 4 (indicated by the yellow
square), next to the CMS experiment. The Timing, Trigger, and Control system
(TTC) is shown as a purple rectangle in the center of the LHC ring. This system
receives RF clocking and broadcasts it into the network (TTC
distribution) through an optical fiber~\cite{ttc_distribution}. The AGM extracts bunch clock 
(corresponding to the position of bunches) and turn clock (marking
the start of a~turn) from the LHC timing block messages.\\
Beam intensity is measured using the synchrotron light (SL) radiation, as indicated by the red line directed towards the Multi-channel photomultiplier
(MCP-PMT) in the optical front-end (OF). The beam traverses first Neutral Density optical filters, which increases the system dynamic range. Subsequently, the signal is
transmitted over a \SI{100}{m}-long cable. To compensate for losses in the cable,
the signal is amplified beforehand. Then it proceeds to the Data Acquisition
(DAQ) system, which measures \SI{6.4}{\micro\second} window and integrates the
intervals of \SI{100}{\nano\second}. This creates set of 64 values published to
the SW every \SI{10}{\micro\second}.
